---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

This is a R Markdown document for Programming Assignment 1 of the Reproducible Research class.  It is written by B. Givens.

## Loading and preprocessing the data

```{r load-data}
# Load data
data <- read.csv("activity.csv")
```

```{r transform-data}
# Process or transform the data for analysis
# Ignore NA
library(dplyr,warn.conflicts=F)
d <- group_by(data,date) %>%
    summarize(day.tot=sum(steps,na.rm=T)) 
# total steps per day
t <- group_by(data,interval) %>%
    summarize(int.avg=mean(steps,na.rm=T)) 
# average steps per interval
```

## What is mean total number of steps taken per day?

```{r histogram, fig.height=4,fig.width=5}
# Make a HISTOGRAM of total steps per day
par(mfrow=c(1,1))
hist(d$day.tot,
     breaks=26,
     xlab="Steps per Day",
     main="Histogram of Steps per Day",
     xlim=c(0,25000),
     ylim=c(0,20),
     col="red")
rug(d$day.tot)
```

```{r mean-median}
# Calculate the mean and median total steps per day
# Ignore NA
d.avg <- mean(d$day.tot)
d.med <- median(d$day.tot)
```

The step average is `r d.avg`.  The median is `r d.med`.

## What is the average daily activity pattern?

```{r time-series-plot,,fig.height=3,fig.width=5}
with(t,plot(interval,int.avg,type="l",
            xlab="Interval",
            ylab="Average Steps",
            main="Daily Activity Pattern",
            col="green",
            lwd=3))
```

```{r find-max}
# Find interval with maximum steps
idx <- which.max(t$int.avg)
t.max <- t$interval[idx]
```

The interval with maximum steps is `r t.max`.

## Imputing missing values

```{r missing-values}
# Count number of missing values
idx <- is.na(data$steps)
count <- sum(idx)
```

The number of NA values is `r count`, which is all intervals for 8 different days.  

```{r devise-strategy}
# Create impute function
# This function will impute the average across dates, or intervals
impute <- function(x) {
    m <- mean(x,na.rm=T)
    if (is.na(m)) {m=0.0}
    bad <- is.na(x)
    x[bad] <- m
    x
}
```

```{r new-dataset}
# Impute NA data across dates
f <- data
dates <- factor(f$date)
temp <- split(f$steps,dates)
temp <- lapply(temp,impute)
f$steps <- unsplit(temp,dates)
```

```{r make-another-hist, fig.height=4,fig.width=5}
# Histogram with imputed data
g <- group_by(f,date) %>%
    summarize(day.tot=sum(steps)) 
# total steps per day
# Make a HISTOGRAM of total steps per day
hist(g$day.tot,
     breaks=26,
     xlab="Steps per Day",
     main="Histogram of Steps per Day, with Imputed Data",
     xlim=c(0,25000),
     ylim=c(0,20),
     col="blue")
rug(g$day.tot)
```

```{r make-another-mean-median}
# Calculate the mean and median total steps per day
# Ignore NA
g.avg <- mean(g$day.tot)
g.med <- median(g$day.tot)
```

The imputed step average is `r g.avg`.  The median is `r g.med`.  

As expected, the histogram, average, and median are same.  Each of 8 days have NAs for all intervals. These NA values are imputed.  The imputed values are zero.  The raw values were assumed zero.  So you get the same thing.

## Are there differences in activity patterns between weekdays and weekends?

```{r create-factor-varible}
# Create weekday/weekend factor variable
days <- weekdays(as.Date(data$date))
wdays <- rep("",length(days))
for (k in seq_along(days)) {
    if (days[k]=="Saturday" | days[k] == "Sunday") {
        wdays[k] <- "weekend"
    }
    else {
        wdays[k] <- "weekday"
    }
}
wdays <- factor(wdays)
```

```{r reduce}
# Reduce data
f$wdays <- wdays
f <- group_by(f,wdays,interval) %>%
    summarize(int.avg=mean(steps)) %>%
    ungroup()
```

```{r lattice}
# Plot the data
library(lattice)
xyplot(int.avg ~ interval | wdays, data=f,
       type="l",
       xlab="Interval",
       ylab="Number of steps",
       layout=c(1,2),
       scales=list(
           y=list(tick.number=5),
           x=list(at=c(0,500,1000,1500,2000))))
```

The activity patterns are different.
